---
layout: tutorial
title: A Quick Introduction to FLR
---

Welcome to FLR. This is going to be a short run through some of the basic elements of the main FLR packages. For further learning material, please visit the [learning page]({{ site.baseurl }}learning/learning.html) or the sites for [each individual package]({{ site.baseurl }}pkgs/pkgs.html).

```{r, ini, echo=FALSE, results='hide', message=FALSE}
render_jekyll()
opts_chunk$set(dev='png', cache=FALSE, fig.width=6, fig.height=6, tidy=TRUE, results='markup')
```

# Installing the FLR packages

We start by installing the most recent version of the FLR packages. Most FLR packages are currently not available in [CRAN](http://cran.r-project.org), but are distributed through the FLR Project package repository. To install the first of the packages, *FLCore*, please type:

```{r}
# install.packages(c('FLCore'),
#  repos='http://flr-project.org/R')
```


# FLCore

```{r, results='hide'}
library(FLCore)
```

```{r, results='hide'}
data(ple4)
plot(ple4)
```
